---
title: "Step by step description of homecredit_wrangling.R"
author: "P. A. Ortiz Otalvaro"
date: "3 September 2019"
output: pdf_document
---

<!---
 describes the variables, the data, and any transformations or work that you performed to clean up the data
 creation date: 21 August 2019
 -->



# 1. Load packages
``` {r results='hide'}
library(dplyr)
library(tidyr)
library(ggplot2)
```


# 2. Read data
For now I am working with the two main files *application_train.csv* and *application_test.csv* in order to define a starting point to improve upon later on.
Data frames `homecredit_train_df` and `homecredit_test_df` contain the data from these two files respectively.

During the reading process, all blank and empty observations are replaced with NA.

```{r}

# *********** Load data (convert any blank or empty entries into NA)*****************
homecredit_train <- readRDS("../LoanData_HomeCredit/application_train.RDS")
homecredit_test <- readRDS("../LoanData_HomeCredit/application_test.RDS")
bureau <- readRDS("../LoanData_HomeCredit/bureau.RDS")
bureaubalance <- readRDS("../LoanData_HomeCredit/bureau_balance.RDS")
previousapp <- readRDS("../LoanData_HomeCredit/previous_application.RDS")
poscashbalance <- readRDS("../LoanData_HomeCredit/POS_CASH_balance.RDS")
installments <- readRDS("../LoanData_HomeCredit/installments_payments.RDS")
creditcard <- readRDS("../LoanData_HomeCredit/credit_card_balance.RDS")

```

<!-----
```{r}
# ******** Add column identifying source (train or test) REMOVE THIS IF I DON'T MERGE THE TWO ########################################################################################################################################################################################################################################################### ********
homecredit_train_df  <- dplyr::mutate(homecredit_train_df , subset = "train")
homecredit_test_df  <- dplyr::mutate(homecredit_test_df , subset = "test")
```
------>



```{r, echo=FALSE}

# ****** get total columns of each data type ******
#train
charcolumns_train <- homecredit_train_df[ , sapply(homecredit_train_df , class) == 'character' ]
factorcolumns_train <- homecredit_train_df[ , sapply(homecredit_train_df, class) == 'factor' ]
numericolumns_train <- homecredit_train_df[ , sapply(homecredit_train_df, class) %in% c('numeric','integer') ]
# test
charcolumns_test <- homecredit_test_df[ , sapply(homecredit_test_df , class) == 'character' ]
factorcolumns_test <- homecredit_test_df[ , sapply(homecredit_test_df, class) == 'factor' ]
numericolumns_test <- homecredit_test_df[ , sapply(homecredit_test_df, class) %in% c('numeric','integer') ]


# ****** get general features of the data set ******
totalrows <- data.frame(t(c(nrow(homecredit_train_df), nrow(homecredit_test_df)) ) , 
                        row.names = "Number of observations")
totalcols <- data.frame(t(c(ncol(homecredit_train_df), ncol(homecredit_test_df)) ) , 
                        row.names = "Number of features")
repeatedrows <- data.frame(t(c(sum(duplicated(homecredit_train_df)) , sum(duplicated(homecredit_test_df)) ) ) , 
                           row.names = "Number of duplicated observations")
missingvalues <- data.frame( t(c( sum(is.na(homecredit_train_df)) , sum(is.na(homecredit_test_df)) )) , 
                             row.names = "Number of missing values" )
nullentries<- data.frame( t(c( sum(is.null(homecredit_train_df)) , sum(is.null(homecredit_test_df)) )) , 
                          row.names = "Number of null entries" )
char_total <- data.frame(t(c( length(charcolumns_train), length(charcolumns_test) ) ) , 
                         row.names = "Total features of char type" )
factor_total <- data.frame(t(c( length(factorcolumns_train), length(factorcolumns_test) ) ), 
                           row.names = "Total features of factor type" )
numeric_total <- data.frame(t(c( length(numericolumns_train), length(numericolumns_test) ) ), 
                            row.names = "Total features of numeric type" )

# ****** all in one data frame ******
general_info_df <- rbind(totalrows, totalcols, repeatedrows, missingvalues, char_total, factor_total, numeric_total)%>%
                   `colnames<-` (c('Train', 'Test'))

# ****** show it as a table ******
knitr::kable(general_info_df)

```
